---
title: "APM1110 FA1 - Navarro"
author: "Angelo M. Navarro"
date: "2026-08-29"
output:
  github_document: default
---



# Pearson's Skewness

<!-- Let's first import the dataset to be used -->
`r results <- read.csv("C:/Users/Angelo/Downloads/results.csv")`

Before we begin, let's have a quick preview of the dataset we are working with:
```{r}
summary(results)
head(results)
```

The fact that there are 120 rows in the dataset makes manual computation incredibly inefficient.
Making an R program would prove to be a far more graceful solution. **Below is a custom function that simulates Pearson's Skewness.**

*Note: There dataset has NA values, so the na.rm argument needs to = TRUE for every stat function used. Otherwise, the said functions would not return a numeric value.*

```{r}
    p.skew <- function(x) {
            3 * (mean(x, na.rm = T) - median(x, na.rm = T)) / sd(x, na.rm = T)
    }
```

\newpage
## Skewness obtained using Pearson's Approximation:

<!-- We get the approximate skewness for each exam and store each in a variable -->
`r pskew.arch1 <- p.skew(results$arch1)`
`r pskew.arch2 <- p.skew(results$arch2)`
`r pskew.prog1 <- p.skew(results$prog1)`
`r pskew.prog2 <- p.skew(results$prog2)`
<!-- Then we create a table displaying the four approximate skewness values -->
`r Skewness_Result <- matrix(c(pskew.arch1, pskew.arch2, pskew.prog1, pskew.prog2), ncol=4)`
`r colnames(Skewness_Result) <- c("ARCH1", "ARCH2", "PROG1", "PROG2")`
`r tab <- as.table(Skewness_Result)`

```{r}
Skewness_Result
```

To check if the modified formula gives a reasonable approximation,
let's compare the answers obtained to how the data is laid out on a histogram.

```{r}
hist(results$arch1, main="ARCH1 (-0.5896811)", xlab= "Exam Scores")
```
\newpage
```{r}
hist(results$arch2, main="ARCH2 (0.5190494)", xlab= "Exam Scores")
```
\newpage
```{r}
hist(results$prog1, main="PROG1 (-0.6345275)", xlab= "Exam Scores")
```
\newpage
```{r}
hist(results$prog2, main="PROG2 (-0.3854051)", xlab= "Exam Scores")
```
\newpage

## Graphical Analysis:
According to the values returned by the **p.skew** function,

+ The **ARCH1, PROG1, and PROG2** exam scores all have **negative skews**; the most pronounced being PROG1 and the least being PROG2. This means more scores are condensed on the right side, thus more students scored above the median.
+ **ARCH2** is the only one with a **positive skew**, which means more students scored below the median.

**All data appears as expected according to the skew function**.
Those with negative skews have most of their data points on the right, and the one with a positive skew has most of its data points on the left.
    
## Exact Skewness
We can also check the accuracy of Pearson's formula by testing it against the original.
**Let's use the syntax provided in Example 2.2:**

```{r}
    e.skew <- function(x) {
              xbar <- mean(x, na.rm = T)
              sum2 <- sum((x-xbar) ^2, na.rm = T)
              sum3 <- sum((x-xbar) ^3, na.rm = T)
              e.skew <- (sqrt(length(x)) * sum3 / (sum2 ^ 1.5))
              }
```

<!-- Like earlier, we get the exact skewness for each exam and store each in a variable -->
`r eskew.arch1 <- e.skew(results$arch1)`
`r eskew.arch2 <- e.skew(results$arch2)`
`r eskew.prog1 <- e.skew(results$prog1)`
`r eskew.prog2 <- e.skew(results$prog2)`
<!-- Then we create another table. This time, displaying the exact skewness values -->
`r Exact_Skewness <- matrix(c(eskew.arch1, eskew.arch2, eskew.prog1, eskew.prog2), ncol=4)`
`r colnames(Exact_Skewness) <- c("ARCH1", "ARCH2", "PROG1", "PROG2")`
`r tab1 <- as.table(Exact_Skewness)`

Now let's check the percent error of each variable:

```{r}
error.arch1 <- (pskew.arch1 - eskew.arch1) / eskew.arch1
error.arch2 <- (pskew.arch2 - eskew.arch2) / eskew.arch2
error.prog1 <- (pskew.prog1 - eskew.prog1) / eskew.prog1
error.prog2 <- (pskew.prog2 - eskew.prog2) / eskew.prog2
```

<!-- Again, we create another table displaying the percent error -->
`r Percent_Error <- matrix(c(error.arch1, error.arch2, error.prog1, error.prog2), ncol=4)`
`r colnames(Percent_Error) <- c("ARCH1", "ARCH2", "PROG1", "PROG2")`
`r tab2 <- as.table(Percent_Error)`

```{r}
Percent_Error
```

## Conclusion
Pearson's formula gives the correct *type* of skewness,
but may give an inaccurate *degree* of skewness.

\newpage

# Displaying Data
First let's input the data from Exercise 1.1:

```{r}
females <- c(57, 59, 78, 79, 60, 65, 68, 71, 75, 48, 51, 55, 56, 41, 43,
44, 75, 78, 80, 81, 83, 83, 85)
males <- c(48, 49, 49, 30, 30, 31, 32, 35, 37, 41, 86, 42, 51, 53, 56,
42, 44, 50, 51, 65, 67, 51, 56, 58, 64, 64, 75)
```

## Stem-and-Leaf Plots
```{r}
stem(females)
stem(males)
```

\newpage
## Advantages of Stem-and-Leaf Plots vs. Histograms:
According to the ChartsMakers Editorial Team (2026)[^1], stem-and-leaf plots are better than histograms for:

* Preserving the original value of data points
* Being simpler to construct (only involves listing and no graphing)
* Representing small data better (n < 50)
* And, stem-and leaf plots can come in back-to-back form for easier comparison between two classes of data. For example (using the aplpack package):

<!--- Let's access the package --->
`r library(aplpack)`

```{r}
stem.leaf.backback(females, males)
```


\newpage
## Box-plots
```{r}
boxplot(females)
```
\newpage
```{r}
boxplot(males)
```

\newpage
## Box-plot Findings

The following categories of findings are based on this article by McLeod (2025)[^2].

**For the females in the class:**

+ The minimum score is slightly above 40
+ The lower quartile ends around the midpoint of 50 and 60
+ The median score is slightly below 70
+ The upper quartile starts slightly below 80
+ The maximum is in the mid-80s
+ The box-plot is negatively skewed

**For the males in the class:**

+ The minimum score is 30
+ The lower quartile ends around 40
+ The median score is 50
+ The upper quartile begins between 50 and 60
+ The upper quartile ends around the mid-70s
+ The maximum value is an outlier in the mid-80s
+ The box-plot is positively skewed

Let's check if our findings are accurate using the summary and exact skew functions on the two vectors:

```{r}
summary(females)
summary(males)
```

```{r}
print(e.skew(females))
print(e.skew(males))
```

## Conclusion:
Our assessment of the box-plot's displayed data is consistent with the actual statistical values.

# References:

+ ChartsMakers. (2026). Stem and Leaf Plot vs Histogram: Key Differences. https://chartsmakers.com/blog/stem-and-leaf-plot-vs-histogram
+ McLeod, S. (2025). Reading a Box and Whisker Plot. SimplyPsychology. https://www.simplypsychology.org/boxplots.html

[^1]: ChartsMakers. (2026). Stem and Leaf Plot vs Histogram: Key Differences.
       https://chartsmakers.com/blog/stem-and-leaf-plot-vs-histogram

[^2]: McLeod, S. (2025). Reading a Box and Whisker Plot. SimplyPsychology.
       https://www.simplypsychology.org/boxplots.html