---
title: "PROB FA3"
author: "Angelo M. Navarro"
date: "2026-09-07"
output:
  pdf_document: default
  html_document: default
---

# Geospatial Analysis System
Here is the dataset we are working with:
<!--Let's set up our data into a table-->
`r Sensors <- c(1, 2, 3, 4)`
`r P_Supplied <- c(0.15, 0.20, 0.25, 0.40)`
`r P_Relevant <- c(0.50, 0.60, 0.80, 0.85)`
`r Geo_Analysis <- matrix(c(Sensors, P_Supplied, P_Relevant), ncol=3)`
`r colnames(Geo_Analysis) <- c("Sensor", "Percentage of Images Supplied", "Percentage of Relevant Images")`
`r tab <- as.table(Geo_Analysis)`

```{r}
Geo_Analysis
```

To get the overall percentage of relevant images, we essentially get the total of each sensor's weighted average. The **weight**, in this case is the **percentage of supplied images**. The elements of the second column are stored in the vector P_Supplied, and those of the third column are stored in P_Relevant.

```{r}
Overall_Relevant <- sum(P_Supplied * P_Relevant) * 100
Overall_Relevant
```

**The overall percentage of relevant images is 73.5%.**

\newpage
# Coin Toss
Let's illustrate the sample space by enumerating all possible combinations of t1 (first toss) and t2 (second toss).
<!--The function below allows us to join the two outcomes into one event-->
```{r}
t1 <- c("H", "T")
t2 <- c("H", "T")
sample.space <- function(..., prefix = "(", suffix=')', sep = ",")
                {paste0(prefix, levels(interaction(..., sep = sep)), suffix)
                }
S <- sample.space(t1, t2)
S
```

Since, this we are using a fair coin, all outcomes have the same probability of occurring (1 in 4 or 0.25). 

Now let's define each event:
```{r}
E1 <- c(S[1], S[4])
E2 <- c(S[1], S[3])
E3 <- c(S[1], S[2])
print(E1)
print(E2)
print(E3)
```

All three events have the same probability at 0.5 (0.25 + 0.25). Let's store them in these variables.
```{r}
PE1 <- 0.5
PE2 <- 0.5
PE3 <- 0.5
```

\newpage
## Pairwise Independence
There are three possible pairings of events. *If the probability of the intersection of two events is equal to the product of their separate probabilities, then they are independent events.*

### Probability of Intersection
The intersect() function returns the common elements of multiple objects.
```{r}
intersect(E1, E2)
intersect(E1, E3)
intersect(E2, E3)
```
All pairings have one event in common, which has a probability of **0.25**.

### Product of Paired Probabilities
```{r}
PE1 * PE2
PE1 * PE3
PE2 * PE3
```

Each pairings intersection probability and product of probability are equal, therefore **E1, E2, and E3 are pairwise independent.**

\newpage
## Mutual Independence
*The three events are mutually independent if the probability the common outcomes of all three are equal to the product of the three events' probability.*

### Probability of Intersection
```{r}
intersect(E1, intersect(E2, E3))
```
There is only one outcome, so probability of their intersection is 0.25.

### Product of Probabilities
```{r}
PE1 * PE2 * PE3
```

**0.25 is not equal to 0.125, therefore they are not mutually independent.**