---
title: "Sec 2 - FA2 Navarro, A - FA2"
author: "Angelo M. Navarro"
date: "2026-08-31"
output: pdf_document
---

## Rolling Dice

```{r}
dice_rolls <- sample(1:6, 600, replace = T)
```

# Frequency per Event
We can count the frequency of each event *(1, 2, 3, 4, 5, 6)* by using the **sum function** and **logical operators**.
The logical operators convert the numerical values into logical values (TRUE or FALSE). **TRUE is equivalent to 1 and FALSE is equivalent to 0.**
```{r}
event_1 <- sum(dice_rolls < 2)
event_2 <- sum(dice_rolls > 1 & dice_rolls < 3)
event_3 <- sum(dice_rolls > 2 & dice_rolls < 4)
event_4 <- sum(dice_rolls > 3 & dice_rolls < 5)
event_5 <- sum(dice_rolls > 4 & dice_rolls < 6)
event_6 <- sum(dice_rolls > 5)
```

To help us visualize the data, let's use a table.
<!-- Syntax for making the table -->
`r Results <- matrix(c(event_1, event_2, event_3, event_4, event_5, event_6), ncol=6)`
`r colnames(Results) <- c("1", "2", "3", "4", "5", "6")`
`r tab <- as.table(Results)`

```{r}
Results
```

\newpage
# Relative Frequencies
To get the relative frequency, we simply need to divide the frequency of each event by the total number of dice rolls (600). **The obtained relative frequency can also serve as the estimated probability for each event.**
```{r}
rfreq_1 <- event_1 / 600
rfreq_2 <- event_2 / 600
rfreq_3 <- event_3 / 600
rfreq_4 <- event_4 / 600
rfreq_5 <- event_5 / 600
rfreq_6 <- event_6 / 600
```

Again, let's use a table to organize our relative frequencies
<!-- Syntax for making the second table -->
`r Estimated_Probabilities <- matrix(c(rfreq_1, rfreq_2, rfreq_3, rfreq_4, rfreq_5, rfreq_6), ncol=6)`
`r colnames(Estimated_Probabilities) <- c("1", "2", "3", "4", "5", "6")`
`r tab <- as.table(Estimated_Probabilities)`

```{r}
Estimated_Probabilities
```