---
title: "PROB FA4"
author: "Angelo M. Navarro"
date: "2026-09-07"
output:
  pdf_document: default
  html_document: default
---

# Binary Communication Channel
Let P(A) denote the probability of a 0 being transmitted and P(B) the probability of 1 being transmitted. P(A) = 0.7.
Then, by the **axiom of certainty and definition of complement**:
```{r}
A <- 0.7
B <- 1 - A
B
```
We can represent these probabilities using a table
<!-- Calibrating table -->
`r Binary_Data <- c(0, 1)`
`r Transmission_Chance <- c(A, B)`
`r Transmissions <- matrix(c(Binary_Data, Transmission_Chance), ncol=2)`
`r colnames(Transmissions) <- c("Binary Data", "Chance of Transmission")`
`r tab <- as.table(Transmissions)`
```{r}
Transmissions
```


Now let P(A1) be the probability of receiving a 0, given that 0 was transmitted. On the other hand, let P(A2) the probability of an error, given that 0 was transmitted. Likewise, B1 is receiving a 1, given that a 1 was transmitted, and B2 is encountering an error while transmitting a 1.
Again, using the **axiom of certainty and definition of complement**:
```{r}
A1 <- 0.95
A2 <- 1 - A1
A2

B1 <- 0.75
B2 <- 1 - B1
B2
```

Let's use a table to organize these probabilities
<!-- Calibrating table -->
`r R_Correct <- c(A1, B1)`
`r R_Error <- c(A2, B2)`
`r Receiving <- matrix(c(R_Correct, R_Error), ncol = 2)`
`r colnames(Receiving) <- c("Received Correctly", "Error in Reception")`
`r tab <- as.table(Receiving)`
```{r}
Receiving
```

## Probability that a 1 was received
Let's represent this event as E1. At first glance this could simply just be P(B1), but an error in transmitting a 0 would also result in receiving a 1. Therefore, E1 is the union of B1 and A2.
So by the **law of total probability** *(with mutually exclusive and exhaustive events)*:
```{r}
E1 = (B * B1) + (A * A2)
```

**The probability of receiving a 1 is 26%.**

## Probability that a 1 was received given that a 1 was transmitted
Let's represent this using P(E2). This is equivalent to the event B1, where a 1 was transmitted and correctly received.
By **law of total probability:**
```{r}
E2 = (B * B1)
```

**The probability of transmitting then receiving a 1 is 7.5%.**

\newpage
# Jane, Amy, and Ava
Let J represent Jane's work, M Amy's, and V Ava's. Let JE, ME, and VE represent their faulty work, respectively.
```{r}
J <- 0.1
M <- 0.3
V <- 0.6

JE <- 0.08
ME <- 0.05
VE <- 0.01
```

Let's tabulate our data for this problem.
<!-- Calibrating table -->
`r P_Program <- c(J, M, V)`
`r P_Error <- c(JE, ME, VE)`
`r IT_Work <- matrix(c(P_Program, P_Error), ncol=2)`
`r colnames(IT_Work) <- c("Proportion of Work Done", "Proportion of Faulty Work")`
`r tab <- as.table(IT_Work)`
```{r}
IT_Work
```

## Overall Percentage of Error
Let's represent this event using TE. This is equivalent to the union of JE, ME, and VE.
Again, using the **law of total probability**:
```{r}
TE <- J * JE + M * ME + V * VE
TE
```

**The overall chance of error is 2.9%.**

## Most likely employee to make a program with error
Let's break down the probability of error by each employee
```{r}
J * JE
M * ME
V * VE
```

**Amy has the highest probability among them all, so she is most likely to make a faulty program.**

